State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-08 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
started working on the discussion

\begin{figure*}[!th] 
\subfigure[][]{\includegraphics[width=2in]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimationError90.pdf}}\\
\caption{One snapshot of the spatial field; a: True field; b: Estimated field; c: The absolute error.}
\label{fig:FieldEstimate}
\end{figure*}
\section{Discussion}\label{DiscussionSection}

Generating patient specific neural field models has the potential to have significant impact on several areas of neuroscientific research and clinical neurology. Specifically, the development of a patient-specific state-space model will allow for the application of systems theoretical techniques from engineering to be applied to neuro-dynamics. For example, state tracking could be used in brain-computer interface or epileptic seizure prediction applications, where specific regions of the state-space may indicate intent of movement or imminent seizures. In addition, a patient specific state-space model will allow application of control engineering techniques to robustly prevent the neuro-dynamics entering pathological regions of state-space using electrical stimulation.

@@

\begin{itemize}
	\item if we don't know the frequency analysis or the system is marginally sampled or under-sampled, the best we can do is have the basis functions the same as the sensor kernel. \wtf{why wouldn't it be better to have, say, twice as many basis functions as we have sensors, at half the width?} 
	\item alternatives to frequency analysis are model selection... \wtf{This frequency analysis \emph{is} model selection. Maybe what's meant here is that other model selection approaches could be applied? Some sort of IC optimisation?} 
	\item the width of the connectivity basis functions \wtf{do you mean width of the kernel? I'm not sure about this - as the parameters increase the kernel width increases right? So maybe you mean the basis functions, but don't we estimate these widths from the frequency analysis?} here is known exactly. In practice we don't 
	\item the number of basis functions could be set higher to get a better representation of the field, however this is limited by computation complexity. \wtf{Unless you are going to specify the computational complexity, in terms of the number of basis functions, it might be better to say something like ``there is a relationship between the number of basis functions that represent the field and the computational complexity of the estimation algorithm - the more basis functions the longer it takes to estimate''} 
	\item Discuss the novelty of using the reduced model (BrainIDE) for creating patient specific models. \omg{hell yes. This should be front and centre!} 
	\item The size of the dentritic arbour is consistent in the cortex. \omg{?} 
	\item Spatial frequency of field and the relationship to the required density of electrodes/sensors required for estimation or to capture the required information without aliasing. 
	\item The flexibility of the approach 
	\item Combining this method with other to incorporate longer range hetrogeneous connectivity. \omg{awesome} 
	\item The need for activity to perform estimation. \omg{To Learn You Must Perturb!} 
	\item Recent studies using neural field models have demonstrated the theoretical implications of specific connectivity structure of the cortical columns in neural field models, where the connectivity kernel governs the bifurcation points of the system \cite{Hutt2005} and types of oscillations that can be generated \cite{Schmidt2009}. This implies that if we could estimate the connectivity structure for an individual, then we could capture patient specific neurodynamics that lead to various oscillatory states.
	\item Estimating functional cortical connectivity via EEG promises to shed light on the hierarchy of brain regions that are involved in specific tasks, motivating the use of techniques like DCM. Other comparable methods are based on auto-regressive (AR) modelling of the EEG, MEG, or fMRI times series \todo{[refs!]}. Studies have concentrated on finding functional connectivity patterns using information contained in the AR coefficients using techniques like Granger causality \cite{Hesse2003}, the direct transfer function \cite{Kaminski1991}, and partial directed coherence \cite{Sameshima1999}. Again, like the DCM approach, these methods estimate long range connectivity patterns that does not provide a clear relationship between the continuum field models and data.
	\item Long range connectivity can be estimated using DTI \cite{Knock2009}, however this short range connectivity can not estimated using DTI
\end{itemize}

>>>>

\begin{figure*}[!th] 
\subfigure[][]{\includegraphics[width=2in]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimationError90.pdf}}\\
\caption{One snapshot of the spatial field; a: True field; b: Estimated field; c: The absolute error.}
\label{fig:FieldEstimate}
\end{figure*}

\section{Discussion}\label{DiscussionSection}
Generating patient specific neural field models has the potential to have significant impact on several areas of neuroscientific research and clinical neurology. Specifically, the development of a patient-specific state-space model will allow for the application of systems theoretical techniques from engineering to be applied to neuro-dynamics. For example, state tracking could be used in brain-computer interface or epileptic seizure prediction applications, where specific regions of the state-space may indicate intent of movement or imminent seizures. In addition, a patient specific state-space model will allow application of control engineering techniques to robustly prevent the neuro-dynamics entering pathological regions of state-space using electrical stimulation.

@@

Testing if this estimation framework can 
We have presented a novel framework for model reduction to make estimation of neural field practical. This systematic model reduction allows for the choice a complexity. The level of detail of the field is related to the number and width of the field basis functions, thereby providing a flexible approach where detail and computational complexity can be tradeoff.. For certain application it might be sufficient to capture only low (spatial) frequency dynamics.

Here we assumed knowledge of the width of the kernel basis functions. If a future work we plan to extend this work to unknown widths.

In future work, we aim to show how heterogeneous can be modelled within this formulation. We will still assume a isotropic connectivity kernel however heterogenaities can be modelled by placing basis function c

\begin{itemize}
	\item if we don't know the frequency analysis or the system is marginally sampled or under-sampled, the best we can do is have the basis functions the same as the sensor kernel. \wtf{why wouldn't it be better to have, say, twice as many basis functions as we have sensors, at half the width?} 
	\item alternatives to frequency analysis are model selection... \wtf{This frequency analysis \emph{is} model selection. Maybe what's meant here is that other model selection approaches could be applied? Some sort of IC optimisation?}

	\item the number of basis functions could be set higher to get a better representation of the field, however this is limited by computation complexity. \wtf{Unless you are going to specify the computational complexity, in terms of the number of basis functions, it might be better to say something like ``there is a relationship between the number of basis functions that represent the field and the computational complexity of the estimation algorithm - the more basis functions the longer it takes to estimate''} 
	\item Discuss the novelty of using the reduced model (BrainIDE) for creating patient specific models. \omg{hell yes. This should be front and centre!} 
	\item The size of the dentritic arbour is consistent in the cortex. \omg{?} 
	\item Spatial frequency of field and the relationship to the required density of electrodes/sensors required for estimation or to capture the required information without aliasing. 
	\item The flexibility of the approach 
	\item Combining this method with other to incorporate longer range hetrogeneous connectivity. \omg{awesome} 
	\item The need for activity to perform estimation. \omg{To Learn You Must Perturb!} 
	\item Recent studies using neural field models have demonstrated the theoretical implications of specific connectivity structure of the cortical columns in neural field models, where the connectivity kernel governs the bifurcation points of the system \cite{Hutt2005} and types of oscillations that can be generated \cite{Schmidt2009}. This implies that if we could estimate the connectivity structure for an individual, then we could capture patient specific neurodynamics that lead to various oscillatory states.
	\item Estimating functional cortical connectivity via EEG promises to shed light on the hierarchy of brain regions that are involved in specific tasks, motivating the use of techniques like DCM. Other comparable methods are based on auto-regressive (AR) modelling of the EEG, MEG, or fMRI times series \todo{[refs!]}. Studies have concentrated on finding functional connectivity patterns using information contained in the AR coefficients using techniques like Granger causality \cite{Hesse2003}, the direct transfer function \cite{Kaminski1991}, and partial directed coherence \cite{Sameshima1999}. Again, like the DCM approach, these methods estimate long range connectivity patterns that does not provide a clear relationship between the continuum field models and data.
	\item Long range connectivity can be estimated using DTI \cite{Knock2009}, however this short range connectivity can not estimated using DTI
\end{itemize}
